\documentclass[../../notes.tex]{subfiles}

\begin{document}

\subsection{Epis and Monos}

\begin{definition}

  In any category C, an arrow $$ f : A \rtarw B $$ is
  called a \textit{monomorphism} if given any $ g, h : C \rtarw A$,
  $gh = fh$ implies $g = h$ \\

  \textit{epimorphism} if given any $i,j : B \rtarw D, if = jf$ implies $i = j$

  % TODO: put in diagrams and arrows with fancy heads/tails

\end{definition}

Remember having a left inverse is monic and having a right inverse is epic. Having both makes the mapping an isomorphism. In SETS, the converse of the previous is also true: every mono-epi is iso; but this is not true in the general case.
% TODO: Propostion 2.6

This definition of monomorphism is the category theory equivalent to injective and this definition of epimorphism is the surjective translation.


\begin{proposition}

  A \textit{function} $ f : A \rtarw B$ between sets is monic just in case it is injective. 
  
\end{proposition}

% \input{initials_terminals}

\begin{definition}{Product}

  In any category $C$, a product diagram for the objects $A$ and $B$
  consists of an object P and arrows
  % \[
  %   \begin{tikzcd}
  %     A & \arrow[l, p_1] P \arrow[r, p_2] & B
  %   \end{tikzcd}
  % \]
  satisfying the universal mapping property: There is some $u: X \rtarw U$
  such that $x_1 = p_1u$ and $x_2 = p_{2}u$. Given any $v: X \rtarw U$, if
  $p_{1}v = x_1$ and $p_{2}v = x_2$ then $v = u$.
\end{definition}

An example:
Let us consider the category of types of the simply typed $\lambda$-calculus.
The $\lambda$ -calculus is a formalism for the specfication and manipulation of functions, based
on the notions of "binding variables" and function evaluation. The relation $ a ~ b $
(usually called $\beta\eta$ -equivalence) on terms is defined to be the equivalence relation
generated by the equations, and the remaining bound variables:
$$ \lambda x.b = \lambda y.b [y/x] (no y in b) $$

The category of types $ C(\lambda ) $ is now defined as follows:
\begin{itemize}
\item Objects: the types
\item Arrows $A \rtarw B$: closed terms $c: A \rtarw B$, identified if $c ~ c'$,
\item Identities $1_A = \lambda x.x (where x : A) $
\item Composition $c \circ b = \lambda x.c(bx)$.
\end{itemize}

\begin{definition}
  A category $C$ is said to have all finite products if it has a terminal object
  and all binary products (and therewith products of any finite cardinality). The
  category $C$ has all (small) products if every set of objects in $C$ has a product.
\end{definition}

\begin{definition}{Slice Category}
  Let $C$ be a category, and $I$ be a $C$-object. Then the category $C / I$, the slice
  category over I, has the following data.
  \begin{itemize}
  \item The objects are pairs (A, f) where $A$ is an object in $C$ and $f: A \rtarw I$ is
    an arrow.
  \item An arrow from $(A, f)$ to $(B, g)$ is an arrow $j: A \rtarw B$ such that $g \circ j = f$ in $C$
  \item The identity arrow on $(A, f)$ is the arrow $1_A : A \rtarw A$.
  \item Given arrows $j: (A, f) \rtarw (B, g)$ and $ k: (B, g) \rtarw (C, h) $, their compostion
    $k \circ j: (A, f) \rtarw (C, h)$ is the arrow $k \circ j : A \rtarw C$.
  \end{itemize}
\end{definition}
\end{document}
